---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at <https://ds202-at-isu.github.io/labs.html> for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.

All submissions to the github repo will be automatically uploaded for grading once the due date is passed. Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
library(tidyverse)
library(dplyr)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`.

-   ***I couldn't run Dhairya Kachalia's code so I wrote my own group part code(Need to comment Dhairya's code to run my code) — Chen***

```         
Completed by Dhairya Kachalia, Chen Sang
```

```{r}
##Dhairya Kachalia's Group Part
#get sum of deaths in Time column
death_cols <-av[c("Death1","Death2","Death3","Death4","Death5")]
return_cols <-av[c("Return1","Return2","Return3","Return4","Return5")]
av <- av %>% mutate(Time = rowSums(death_cols == "YES",na.rm = TRUE))
av <- av %>% mutate(Return_Time = rowSums(return_cols == "YES",na.rm = TRUE))

#remove Death(2-5) and Return(2-5)
av <- av[,-(13:20)]

#rename Death1 to Death and Return1 to Return
av<-av %>% rename("Death"="Death1")
av<-av %>% rename("Return"="Return1")

rm(death_cols)


###Chen's Part for group part

# Perform transformations and store the result in 'av_transformed'
av_transformed <- av %>%
  mutate(
    Time = rowSums(select(., starts_with("Death")) == "YES", na.rm = TRUE),
    Return_Time = rowSums(select(., starts_with("Return")) == "YES", na.rm = TRUE)
  ) %>%
  select(-matches("^Death[2-5]$"), -matches("^Return[2-5]$")) %>%
  rename(Death = Death1, Return = Return1)

# View the modified dataframe
print(av_transformed)

```

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers.

```{r}
##Dhairya Kachalia's Group Part
cat("The average deaths an avenger get is ", mean(av$Time),"\n")


###Chen's part
# Calculate the average number of deaths
average_deaths <- mean(av_transformed$Time, na.rm = TRUE)

# Print the average number of deaths
print(average_deaths)
```

**Chen's part: I couldn't run Dhairya Kachalia's code So I have my own code for group part**

-   Need to comment out Dhairya Kachalia's code to be able to run my code

```         
```

## Individually

For each team member, copy this part of the report.

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

```         
Dhairya Kachalia analysis statement:
Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team. That’s about 40 percent of all people who have ever signed on to the team. 
```

```{r}
#get rows where Time >=1
filter_row <- av[av$Time >=1,]
cat("the total number of avengers that died atleast once is", nrow(filter_row), "\n")
```

```         
statement2: Of the nine Avengers we see on screen — Iron Man, Hulk, Captain America, Thor, Hawkeye, Black Widow, Scarlet Witch, Quicksilver and The Vision — every single one of them has died at least once in the course of their time Avenging in the comics. In fact, Hawkeye died twice!
```

```{r}
# i will use the names Anthony Edward "Tony" Stark, Robert Bruce Banner, Thor Odinson, Steven Rogers, Pietro Maximoff, Wanda Maximoff, Natalia Alianovna Romanova, Clinton Francis Barton instead of Iron Man, Hulk, Hawkeye, etc. as these are the names in the database

#gets the rows where name alias are as above and Time >=1
tv_avengers <- av %>% select(Name.Alias,Time) %>%  filter(Name.Alias %in% c("Anthony Edward \"Tony\" Stark", "Robert Bruce Banner", "Thor Odinson", "Steven Rogers", "Pietro Maximoff", "Wanda Maximoff", "Natalia Alianovna Romanova", "Clinton Francis Barton"),av$Time >=1)

tv_avengers
# the tv_anvegers table shows that the avengers in this analysis died atleast once and Hawkeye i.e Clinton Francis Barton died twice
```

```         
statement 3: I counted 89 total deaths — some unlucky Avengers7 are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback. But you can only tempt death so many times. There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.
```

```{r}

#gives the sum of the total number of deaths in av. Also adds multiple deaths of same character
sum_deaths <- sum(av$Time, na.rm = TRUE)

#gives the sum of the total number of returns in av. Also adds multiple deaths of same character
sum_returns <- sum(av$Return_Time, na.rm = TRUE)

cat("Total occasions of deaths is",sum_deaths,"which matches the statement.\n")
cat("Total ocassions of returns is", sum_returns,"which matches the statement\n")

#gives the sum of total characters who have died atleast once
sum_died_atleast_once <- nrow(av %>% filter(av$Time >=1))

#gives the sum of total characters who have made a comeback atleast once
sum_returns_atleast_once <- nrow(av %>% filter(av$Return_Time>=1))

cat("The percentage of characters who made a comeback atleast once is",(sum_returns_atleast_once/sum_died_atleast_once)*100,"% which matches the statement that the comeback rate is 2 in 3 \n")

#characters who made more than 1 return
gg <- av %>% filter(av$Return_Time>1)
#sum of ocassions where the character made more than one time
sum_returns_more_than_once <-sum(gg$Return_Time, na.rm = TRUE)

```

```         
```

# Chen Sang's Part:

### FiveThirtyEight Statement

For each team member, copy this part of the report.

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible. \`\`\`

-   Chen Sang's Statement:

    -   "There’s a 2/3 chance that a member of the Avengers returned from their first stint in the afterlife."

        ```{r}
        # Correct usage of filter without data frame prefix inside the function
        total_deaths_occasions <- av_transformed %>%
          filter(Time >= 1) %>%
          nrow()

        # Similarly, for total occasions of returns if you have a 'Return_Time' or similar column
        total_returns_occasions <- av_transformed %>%
          filter(Return_Time >= 1) %>%
          nrow()

        # Print the results
        print(paste("Total occasions of deaths:", total_deaths_occasions)) #69 death
        print(paste("Total occasions of returns:", total_returns_occasions)) # 46 return
        ```

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.
